---
title: "work_normalization-etc_rough-draft_NNS_vary-on-transcription"
author: "KA"
email: "kalavatt@fredhutch.org"
output: html_notebook
---
<br />

## Prepare data for various DGE analyses
### Get situated
#### Set working directory
```{r Set working directory, results='hide', message=FALSE}
# p_local <- "/Users/kalavattam/Dropbox/FHCC"  # KrisMac
p_local <- "/Users/kalavatt/projects-etc"  # WorkMac
p_wd <- "2022_transcriptome-construction/results/2023-0215"
#NOTE 1/3 Alison: you can adjust 'p_local' or 'p_wd' (path to working directory 
#NOTE 2/3 on your personal or work computer ); I have two paths here with one
#NOTE 3/3 commented out depending on which computer I am using

setwd(paste(p_local, p_wd, sep = "/"))
getwd()

rm(p_local, p_wd)
```
<br />

#### Load necessary libraries
```{r Load necessary libraries, results='hide', message=FALSE, warning=FALSE}
library(DESeq2)
library(EnhancedVolcano)
library(GenomicRanges)
library(IRanges)
library(tidyverse)
library(readxl)  # Needed for reading in Excel file of filenames, variables
```
<br />

#### Use normal numbers instead of default scientific numbers in plots
The `scipen` option determines how likely R is to switch to scientific notation;
the higher the value, the less likely it is to switch. We're setting the option
prior to making plots; if plots still have scientific notation, then we need to
set the option to a higher number.

```{r}
options(scipen=999)
```
<br />

#### Initialize necessary functions
```{r Initialize necessary functions, results='hide', message=FALSE}
split_isolate_convert <- function(in_vector, field, column_name) {
    # :param in_vector: ...
    # :param field: ...
    # :param column_name: ...
    # :return out_df: ...
    out_df <- in_vector %>%
        stringr::str_split(., c("_")) %>%
        sapply(., "[", field) %>%
        as.data.frame() %>%
        tibble::as_tibble()
    
    colnames(out_df) <- column_name
    
    return(out_df)
}
#TODO Note where this function is being used
#TODO Add description of what function does
#TODO Add parameter and return descriptions


plot_volcano <- function(
    table, label, selection, label_size, p_cutoff, FC_cutoff,
    xlim, ylim, color, title, subtitle, ...
) {
    # :param table: dataframe of test statistics
    # :param label: character vector of all variable names in param table
    # :param selection: character vector of selected variable names in param
    #                   table
    # :param label_size: size of label font [float]
    # :param p_cutoff: cut-off for statistical significance; a horizontal line
    #                  will be drawn at -log10(pCutoff); p is actually padj
    # :param FC_cutoff: cut-off for absolute log2 fold-change; vertical lines
    #                   will be drawn at the negative and positive values of
    #                   log2FCcutoff
    # :param xlim: limits of the x-axis
    # :param ylim: limits of the y-axis
    # :param color: color of DEGs, e.g., '#52BE9B'
    # :param title: plot title
    # :param subtitle: plot subtitle
    # :return volcano: ...
    volcano <- EnhancedVolcano::EnhancedVolcano(
        toptable = table,
        lab = label,
        selectLab = selection,
        x = "log2FoldChange",
        y = "padj",
        xlab = "log2(FC)",
        ylab = "-log10(padj)",
        pCutoff = p_cutoff,
        pCutoffCol = "padj",
        FCcutoff = FC_cutoff,
        xlim = xlim,
        ylim = ylim,
        cutoffLineType = "dashed",
        cutoffLineWidth = 0.2,
        pointSize = 1,
        shape = 16,
        colAlpha = 0.25,
        col = c('#D3D3D3', '#D3D3D3', '#D3D3D3', color),
        title = NULL,
        subtitle = NULL,
        caption = NULL,
        borderColour = "#000000",
        borderWidth = 0.2,
        gridlines.major = TRUE,
        gridlines.minor = TRUE,
        axisLabSize = 10,
        labSize = label_size,
        boxedLabels = TRUE,
        parseLabels = TRUE,
        drawConnectors = TRUE,
        widthConnectors = 0.2,
        colConnectors = 'black',
        max.overlaps = Inf
    ) +
        theme_slick_no_legend +
        ggplot2::ggtitle(title, subtitle = subtitle)
    return(volcano)
}
#TODO Note where this function is being used
#TODO Add description of what function does
#TODO Add return description


save_volcano <- function(plot, file, width, height) {
    # :param plot: ...
    # :param file: ...
    # :param width: ...
    # :param height: ...
    # :return: ...
    ggplot2::ggsave(
        plot,
        filename = file,
        device = "pdf",
        h = width,
        w = height,
        units = "in"
    )
}
#TODO Note where this function is being used
#TODO Add description of what function does
#TODO Add return description


#  Set up custom ggplot2 plot themes ------------------------------------------
theme_slick <- theme_classic() +
    theme(
        # panel.grid.major = ggplot2::element_line(linewidth = 0.4),
        # panel.grid.minor = ggplot2::element_line(linewidth = 0.2),
        panel.grid.major = ggplot2::element_line(size = 0.4),
        panel.grid.minor = ggplot2::element_line(size = 0.2),
        # axis.line = ggplot2::element_line(linewidth = 0.2),
        # axis.ticks = ggplot2::element_line(linewidth = 0.4),
        axis.line = ggplot2::element_line(size = 0.2),
        axis.ticks = ggplot2::element_line(size = 0.4),
        axis.text = ggplot2::element_text(color = "black"),
        axis.title.x = ggplot2::element_text(),
        axis.title.y = ggplot2::element_text(),
        plot.title = ggplot2::element_text(),
        text = element_text(family = "")
    )
#TODO Note where this theme is being used
#TODO 1/2 Add explanation of why there are lines where 'size' is used and
#TODO 2/2 alternatives where 'linewidth' is used

theme_slick_no_legend <- theme_slick + theme(legend.position = "none")
#TODO Note where this theme is being used
```
<br />

### Load in Excel spreadsheet of samples names and variables
The spreadsheet includes Alison's original sample names; we can use this
information to associate the new sample names, which are made up of `DESeq2`
GLM variables, with the old names, which reflect Alison's wet-lab, library-prep,
etc. work
```{r load spreadsheet, results='hide', message=FALSE}
p_xl <- "notebook"  #INPATH
f_xl <- "variables.xlsx"  #INFILE
#NOTE 1/4 Alison: you can adjust 'p_xc' (path for Excel file) and "f_xl" (the
#NOTE 2/4 Excel file proper) to point to wherever UT_prim_UMI.featureCounts is
#NOTE 3/4 on your system; "t_xl" stands for, more or less, "tibble for the
#NOTE 4/4 Excel file"
t_xl <- readxl::read_xlsx(
    paste(p_xl, f_xl, sep = "/"), sheet = "master", na = "NA"
)

rm(p_xl, f_xl)
```
<br />

### Load in and process `featureCounts` table
```{r Process featureCounts table, results='hide', message=FALSE}
#  Load in featureCounts table ------------------------------------------------
f_fc <- "UT_prim_UMI.featureCounts"  #INFILE
#NOTE 1/2 Alison: you can adjust 'p_fc' (path for featureCounts file) and
#NOTE 2/3 'f_fc' (featureCounts file) to point to wherever
#NOTE 3/3 "UT_prim_UMI.featureCounts" is on your system; 't_fc' stands for,
#NOTE 4/4 "tibble for featureCounts file" (or something like that)
t_fc <- read.table(
    paste(p_fc, f_fc, sep = "/"), header = TRUE, row.names = 1
) %>% 
    tibble::rownames_to_column() %>%
    tibble::as_tibble()

rm(p_fc, f_fc)


#  Clean up tibble column names -----------------------------------------------
colnames(t_fc) <- colnames(t_fc) %>%
    gsub("rowname", "feature_init", .) %>%
    gsub("Chr", "chr", .) %>%
    gsub("Start", "start", .) %>%
    gsub("End", "end", .) %>%
    gsub("Strand", "strand", .) %>%
    gsub("Length", "length", .) %>%
    gsub("bams_renamed\\.UT_prim_UMI\\.", "", .) %>%
    gsub("\\.UT_prim_UMI\\.bam", "", .) %>%
    gsub("\\.d", "-d", .) %>%
    gsub("\\.n", "-n", .) %>%
    gsub("aux\\.", "aux-", .) %>%
    gsub("tc\\.", "tc-", .)


#  Order tibble by chromosome names and feature start positions ---------------
chr_SC <- c(
    "I", "II", "III", "IV", "V", "VI", "VII", "VIII", "IX", "X", "XI", "XII",
    "XIII", "XIV", "XV", "XVI", "Mito"
)
chr_KL <- c("A", "B", "C", "D", "E", "F")
chr_order <- c(chr_SC, chr_KL)
t_fc$chr <- t_fc$chr %>% as.factor()
t_fc$chr <- ordered(t_fc$chr, levels = chr_order)

t_fc <- t_fc %>% dplyr::arrange(chr, start)


#  Categorize chromosomes by genome of origin ---------------------------------
t_fc$genome <- ifelse(
    t_fc$chr %in% chr_SC,
    "S_cerevisiae",
    ifelse(
        t_fc$chr %in% chr_KL,
        "K_lactis",
        NA
    )
) %>%
    as.factor()

#  Move the new column "genome" to a better location in the tibble (before
#+ column "chr")
t_fc <- t_fc %>% dplyr::relocate("genome", .before = "chr")

#  Check on variable/column "genome"
levels(t_fc$genome)
t_fc %>%
    dplyr::group_by(genome) %>%
    dplyr::summarize(tally = length(genome))
#  The code returns...
# K_lactis = 5659, S_cerevisiae = 7507

rm(chr_KL, chr_SC, chr_order)


#  Split and better organize variable 'feature_init' --------------------------
#  Split 'feature_init' into two distinct elements (separated by an underscore)
el_1 <- split_isolate_convert(
    in_vector = t_fc$feature_init,
    field = 1,
    column_name = "feature"
)
el_2 <- split_isolate_convert(
    in_vector = t_fc$feature_init,
    field = 2,
    column_name = "type"
)

#  Append split information to tibble 't_fc'
t_fc <- dplyr::bind_cols(t_fc, el_1, el_2) %>%
    dplyr::relocate(c("feature", "type"), .after = "feature_init")

rm(el_1, el_2)

#  Limit the splitting/reorganization to S. cerevisiae features only; the above
#+ splitting/reorganization work isn't appropriate for K. lactis 'feature_init'
#+ information because the K. lactis naming/classification differs from the S.
#+ cerevisiae naming/classification system
t_fc$feature <- ifelse(
    t_fc$genome == "K_lactis", t_fc$feature_init, t_fc$feature
)
t_fc$type <- ifelse(
    t_fc$genome == "K_lactis", NA, t_fc$type
)

#  Create levels for S. cerevisiae 'type' NAs and K. lactis 'type' NAs, then
#+ factorize variable 'type': essentially, we're making the NAs into levels so
#+ that we can tally them (as below) and/or potentially subset them; however,
#+ before doing so, we're differentiating the NAs by whether they are
#+ associated with S. cerevisiae features or K. lactis features
t_fc$type <-  ifelse(
    (t_fc$genome == "S_cerevisiae" & is.na(t_fc$type)),
    "NA_SC",
    ifelse(
        (t_fc$genome == "K_lactis" & is.na(t_fc$type)),
        "NA_KL",
        t_fc$type
    )
) %>%
    as.factor()

#  Do a quick check of the tibble 't_fc' (where "t_fc" stands for "tibble
#+ featureCounts")
t_fc

#  Check on the split information: This code tallies the numbers features per
#+ classification, where classifications are things like "mRNA-E1", "tRNA-E1",
#+ "NA_SC" (NAs associated with S. cerevisiae), "NA_KL" (NAs associated with K.
#+ lactis), etc.
levels(t_fc$type)  # 19 levels
t_fc %>%
    dplyr::group_by(type) %>%
    dplyr::summarize(tally = length(type))
#  The code returns things like...
#+ mRNA-E1 = 6600, mRNA-E2 = 283, NA_KL = 5547, NA_SC = 103, tRNA-E1 = 299,
#+ tRNA-E2 = 60, etc.
```
<br />

### Record tibble `t_fc`'s positional information in a `GRanges` object
`pos_info` will be used in `DESeq2` processing, post-processing, etc.

```{r Record positional information, results='hide', message=FALSE}
pos_info <- GenomicRanges::GRanges(
    seqnames = t_fc$chr,
    ranges = IRanges::IRanges(t_fc$start, t_fc$end),
    strand = t_fc$strand,
    length = t_fc$length,
    feature = t_fc$feature,
    type = t_fc$type,
    genome = t_fc$genome
)
pos_info
```
<br />
<br />

## Perform normalization and run DGE analyses
### Perform prep work
#### Establish table of variables for `dds`&mdash;i.e., a "master" model matrix
- `dds` stands for *"DESeq2 dataset"* and is a `DESeqDataSet` object
- variables for `dds` are
    + `strain`
    + `state`
    + `time`
    + `kit` *(`tcn` for "Tecan", `ovn` for "Ovation")*
    + `transcription` *(`N` for "nascent", `SS` for "steady state")*
    + `auxin`
    + `timecourse`
    + `replicate`

```{r Make a master model matrix, results='hide', message=FALSE}
#  Columns ten through to the last column are composed of sample feature
#+ counts; get these column names into a vector
samples <- colnames(t_fc)[10:length(colnames(t_fc))]

#  Convert the vector of column names to a list by splitting each element at
#+ its underscores; thus, each vector element becomes a list of eight strings,
#+ with one string for 'strain', one for 'state', etc.; these 
samples <- stringr::str_split(samples, "_")

#  Convert the list to a df, transpose it, then convert it to a tibble
#+ (R fun fact: 'tibble' data types can't be built directly from list data
#+ types)
samples <- samples %>%
    as.data.frame(
        .,
        #  Using numeric column names here because the columns will soon be
        #+ transposed to rows, and I don't want the rows to have proper names
        col.names = c(seq(1, 55)),
        #  Using proper row names here because the rows will soon be transposed
        #+ to columns, and I *do* want the columns to have proper names 
        row.names = c(
            "strain", "state", "time", "kit", "transcription", "auxin",
            "timecourse", "replicate"
        )
    ) %>%
    t() %>%
    tibble::as_tibble()

#  Add a keys variable for quickly accessing combinations of variable values
keys <- vector(mode = "character")
for(i in seq(1, nrow(samples))) {
    # i <- 1
    keys[i] <- paste(
        samples[i, 1], samples[i, 2], samples[i, 3], samples[i, 4],
        samples[i, 5], samples[i, 6], samples[i, 7], samples[i, 8],
        sep = "_"
    )
}
keys <- keys %>% as.data.frame()
colnames(keys) <- "keys"

samples <- dplyr::bind_cols(samples, keys) %>%
    dplyr::relocate("keys", .before = "strain")

rm(i)

#  Add Alison's original samples names to the 'samples' dataframe using the
#+ 't_xl' dataframe; here, we're just adding the original sample names, but we
#+ could potentially add in other information stored in the Excel file
t_xl <- t_xl %>%
    dplyr::rename(keys = name) %>%
    dplyr::select(., c(keys, sample_name))
samples <- dplyr::full_join(samples, t_xl, by = "keys")

#  Convert all columns to data type 'factor' (having the variable values as
#+ factors helps with running DESeq2::DESeqDataSetFromMatrix() below)
samples[sapply(samples, is.character)] <- lapply(
    samples[sapply(samples, is.character)], as.factor
)

#  How does it look?
samples

rm(t_xl, keys)
```
<br />

### Do prep work with `WT_{G1,Q}_day7_ovn_N_aux-F_tc-F_rep{1,2}`
That is, do prep work with the wild-type G1 and Q Ovation nascent
(immunoprecipitated) 4tU-seq datasets, replicates 1 and 2.
<br />

#### Make the counts matrix
```{r Make the counts matrix, results='hide', message=FALSE}
datasets <- c(
    "n3-d_Q_day7_tcn_N_aux-T_tc-F_rep1",
    "n3-d_Q_day7_tcn_N_aux-T_tc-F_rep2",
    "n3-d_Q_day7_tcn_N_aux-T_tc-F_rep3",
    "n3-d_Q_day7_tcn_SS_aux-T_tc-F_rep1",
    "n3-d_Q_day7_tcn_SS_aux-T_tc-F_rep2",
    "n3-d_Q_day7_tcn_SS_aux-T_tc-F_rep3"
)
counts_data <- t_fc[, colnames(t_fc) %in% datasets] %>%
    as.data.frame()  #IMPORTANT Output a dataframe, not a tibble

#  How do things look?
counts_data
```
<br />

#### Make the model matrix
```{r make model matrix, results='hide', message=FALSE}
#  Use the "keys" column to isolate datasets of interest
#REMEMBER Variable `datasets` is initialized in the preceding chunk 
col_data <- samples[samples$keys %in% datasets, ] %>%
    as.data.frame() %>%    #IMPORTANT Output a dataframe, not a tibble
    tibble::column_to_rownames(., var = "keys")  #IMPORTANT Have row names

#  How do things look?
col_data
```
<br />

#### Make the `DESeqDataSet`, `dds`
- Use `counts_data` for the `featureCount` tallies
- Use `col_data` for setting up the GLM
- Use `pos_info` for adding feature metadata, subsequent subsetting, etc.

```{r make dds, results='hide', message=FALSE}
dds <- DESeq2::DESeqDataSetFromMatrix(
    countData = counts_data,
    colData = col_data,
    design = ~ transcription,
    rowRanges = pos_info
)

#  Make a back-up of the DESeqDataSet object
bak.dds <- dds

#  How do things look?
# dds %>% BiocGenerics::counts() %>% head()
# dds@rowRanges
# dds@design
# dds@assays
```
<br />

#### Prefilter `dds`
We probably don't need to do this, but then again we may want to; `#TODO` Let's
make a decision later.

```{r prefilter dds, results='hide', message=FALSE}
# threshold <- 0
# dds_filt <- dds[rowSums(BiocGenerics::counts(dds)) >= threshold, ]
#
#  Breakdown
#    0 13166
#   10 11893
#   50 11049
#  100 10645
#  500 8601
# 1000 6738
#
# rm(threshold, dds_filt)
```
<br />
<br />

## I. Run analyses with *S. cerevisae* features only
### Perform size-factor estimation
Here, we subset out the *K. lactis* features. Thus, we are using only
*S. cerevisiae* features in the size-factor estimation. No control genes are
used.

```{r I. Perform size-factor estimation, message=FALSE}
dds_SC <- BiocGenerics::estimateSizeFactors(
    dds[dds@rowRanges$genome != "K_lactis", ]
)
dds_SC@colData
# Q N rep1 1.914013
# Q N rep2 1.489813
# Q N rep3 2.061973
# Q SS rep1 0.695266
# Q SS rep2 0.497825
# Q SS rep3 0.505492
```
<br />

### Run `DESeq2`
#### Call `DESeq2` using default parameters
```{r I. Call DESeq2 using default parameters, results='hide', message=FALSE}
dds_SC <- DESeq2::DESeq(dds_SC)

#  Check model information
DESeq2::resultsNames(dds_SC)[2]
#  [1] "transcription_SS_vs_N"
#+ Thus, the model varies on transcription, SS is the numerator, N is the
#+ denominator
```
<br />

#### Call `DESeq2::results()`
```{r I. Call DESeq2::results(), message=FALSE}
#  Set up necessary parameters for generation of DESeq2 results table
independent_filtering <- TRUE
threshold_p <- 0.05
threshold_lfc <- 0

#  Output a DESeq2 DataFrame object
DGE_unshrunken_DF_SC <- DESeq2::results(
    dds_SC,
    name = DESeq2::resultsNames(dds_SC)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "DataFrame"
)

#  Output a GRanges object, which we can easily add to and convert to other
#+ formats (such as a tibble)
DGE_unshrunken_GR_SC <- DESeq2::results(
    dds_SC,
    name = DESeq2::resultsNames(dds_SC)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "GRanges"
)
DGE_unshrunken_GR_SC$feature <- MatrixGenerics::rowRanges(dds_SC)$feature
DGE_unshrunken_GR_SC$type <- MatrixGenerics::rowRanges(dds_SC)$type
DGE_unshrunken_GR_SC$genome <- MatrixGenerics::rowRanges(dds_SC)$genome

t_DGE_SC <- DGE_unshrunken_GR_SC %>%
    dplyr::as_tibble() %>%
    dplyr::rename(chr = seqnames)

rm(independent_filtering, threshold_p, threshold_lfc)
```
<br />

#### Make an MA plot that colors features by independent filtering
```{r I. Make an MA plot: B, message=FALSE}
#  Set up temporary variable 'tbl', which will be passed to ggplot
tbl <- t_DGE_SC
tbl <- tbl[with(tbl, order(log2FoldChange)), ]
tbl$threshold <- as.factor(tbl$padj <= 0.05)
tbl$log10baseMean <- ifelse(
    is.infinite(log10(tbl$baseMean)), NA, log10(tbl$baseMean)
)

title <- paste0(
    "MA plot | S. cerevisiae features |\n",
    "size factors estimated with all S. cerevisiae features"
)
subtitle <- paste(
    "points: S. cerevisiae features",
    "| top: up in SS",
    "| bottom: up in N"
)  # [1] "transcription_SS_vs_N"
ggplot(tbl, aes(x = log10baseMean, y = log2FoldChange, colour = threshold)) +
    geom_point(alpha = 0.25, size = 0.5) +
    # geom_hline(aes(yintercept = 0), colour = "#000000", linewidth = 0.25) +
    geom_hline(aes(yintercept = 0), colour = "#000000", size = 0.25) +
    # ylim(c(min(tbl$log2FoldChange), max(tbl$log2FoldChange))) +
    ylim(c(-14, 14)) +
    xlab("log10(mean normalized counts)") +
    ylab("log2(fold change)") +
    scale_colour_discrete(name = "q ≤ 0.05") +
    ggtitle(title, subtitle) +
    theme_slick
#TODO 1/2 Explain and make a decision regarding use of 'size' or 'linewidth'
#TODO 2/2 parameters

#  Create a vector of features that both passed independent filtering (and thus
#+ have an inherently high mean expression) and are not statistically
#+ significant; this vector signifies features that are "stably expressed"
#+ between conditions
tbl$stably_expressed <- ifelse(
    !is.na(tbl$threshold) & tbl$padj > 0.05,
    TRUE,
    FALSE
)
stably_expressed_SC <- tbl$feature[tbl$stably_expressed == TRUE]

rm(tbl, title, subtitle)
```
<br />

#### Make a volcano plot
```{r I. Make a volcano plot, message=FALSE}
#  Identify and isolate the top 5 upregulated features and the top 5 down-
#+ regulated features
all <- t_DGE_SC$feature
selection_down <- t_DGE_SC %>%
    dplyr::filter(log2FoldChange < 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection_up <- t_DGE_SC %>%
    dplyr::filter(log2FoldChange > 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection <- c(selection_down[["feature"]], selection_up[["feature"]]) %>%
        as.character()

title <- paste0(
    "volcano plot | S. cerevisiae features |\n",
    "size factors estimated with all S. cerevisiae features"
)
subtitle <- paste(
    "points: S. cerevisiae features",
    "| left: up in N",
    "| right: up in SS",
    "| labels: top 5 N and top 5 SS features"
)  # [1] "transcription_SS_vs_N"
p_SC <- plot_volcano(
    table = t_DGE_SC,
    label = all,
    selection = selection,
    label_size = 2.5,
    p_cutoff = 0.05,
    FC_cutoff = 1,
    xlim = c(-14, 14),
    ylim = c(0, 310),
    color = "#52BE9B",
    title = title,
    subtitle = subtitle
)
p_SC
# save_volcano(
#     file = "test.pdf",
#     plot = p,
#     width = 2,
#     height = 3
# )

rm(all, selection, selection_up, selection_down, title, subtitle)
```
<br />
<br />

## II. Run analyses with *K. lactis* features only
### Perform size-factor estimation
Here, we subset out&mdash;i.e., remove&mdash;the *S. cerevisiae* features and
are thus only analyzing *K. lactis* features, using all *K. lactis* features
in the size-factor estimation.

```{r II. Perform size-factor estimation, results='hide', message=FALSE}
dds_KL <- BiocGenerics::estimateSizeFactors(
    dds[dds@rowRanges$genome != "S_cerevisiae", ]
)
dds_KL@colData
# Q N rep1 6.264768
# Q N rep2 7.190513
# Q N rep3 6.525544
# Q SS rep1 0.161401
# Q SS rep2 0.165265
# Q SS rep3 0.140550
```
<br />

### Run `DESeq2`
#### Call `DESeq2` using default parameters
```{r II. Call DESeq2 using default parameters, results='hide', message=FALSE}
dds_KL <- DESeq2::DESeq(dds_KL)

#  Check model information
DESeq2::resultsNames(dds_KL)[2]
#  [1] "transcription_SS_vs_N"
#+ Thus, the model varies on transcription, SS is the numerator, N is the
#+ denominator
```
<br />

#### Call `DESeq2::results()`
```{r II. Call DESeq2::results(), message=FALSE}
#  Set up necessary parameters for generation of DESeq2 results table
independent_filtering <- TRUE
threshold_p <- 0.05
threshold_lfc <- 0

#  Output a DESeq2 DataFrame object
DGE_unshrunken_DF_KL <- DESeq2::results(
    dds_KL,
    name = DESeq2::resultsNames(dds_KL)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "DataFrame"
)

#  Output a GRanges object, which we can easily add to and convert to other
#+ formats (such as a tibble)
DGE_unshrunken_GR_KL <- DESeq2::results(
    dds_KL,
    name = DESeq2::resultsNames(dds_KL)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "GRanges"
)
DGE_unshrunken_GR_KL$feature <- MatrixGenerics::rowRanges(dds_KL)$feature
DGE_unshrunken_GR_KL$type <- MatrixGenerics::rowRanges(dds_KL)$type
DGE_unshrunken_GR_KL$genome <- MatrixGenerics::rowRanges(dds_KL)$genome

t_DGE_KL <- DGE_unshrunken_GR_KL %>%
    dplyr::as_tibble() %>%
    dplyr::rename(chr = seqnames)

rm(independent_filtering, threshold_p, threshold_lfc)
```
<br />

#### Make an MA plot that colors features by independent filtering
```{r II. Make an MA plot: B, message=FALSE}
#  Set up temporary variable 'tbl', which will be passed to ggplot
tbl <- t_DGE_KL
tbl <- tbl[with(tbl, order(log2FoldChange)), ]
tbl$threshold <- as.factor(tbl$padj <= 0.05)
tbl$log10baseMean <- ifelse(
    is.infinite(log10(tbl$baseMean)), NA, log10(tbl$baseMean)
)

title <- paste0(
    "MA plot | K. lactis features |\n",
    "size factors estimated with all K. lactis features"
)
subtitle <- paste(
    "points: K. lactis features",
    "| top: up in SS",
    "| bottom: up in N"
)  # [1] "transcription_SS_vs_N"
ggplot(tbl, aes(x = log10baseMean, y = log2FoldChange, colour = threshold)) +
    geom_point(alpha = 0.25, size = 0.5) +
    # geom_hline(aes(yintercept = 0), colour = "#000000", linewidth = 0.25) +
    geom_hline(aes(yintercept = 0), colour = "#000000", size = 0.25) +
    # ylim(c(min(tbl$log2FoldChange), max(tbl$log2FoldChange))) +
    ylim(c(-14, 14)) +
    xlab("log10(mean normalized counts)") +
    ylab("log2(fold change)") +
    scale_colour_discrete(name = "q ≤ 0.05") +
    ggtitle(title, subtitle) +
    theme_slick
#TODO 1/2 Explain and make a decision regarding use of 'size' or 'linewidth'
#TODO 2/2 parameters

#  Create a vector of features that both passed independent filtering (and thus
#+ have an inherently high mean expression) and are not statistically
#+ significant; this vector signifies features that are "stably expressed"
#+ between conditions
tbl$stably_expressed <- ifelse(
    !is.na(tbl$threshold) & tbl$padj > 0.05,
    TRUE,
    FALSE
)
stably_expressed_KL <- tbl$feature[tbl$stably_expressed == TRUE]
#NOTE Probably need to be more stringent here

rm(tbl, title, subtitle)
```
<br />

#### Make a volcano plot
```{r II. Make a volcano plot, message=FALSE}
all <- t_DGE_KL$feature
selection_down <- t_DGE_KL %>%
    dplyr::filter(log2FoldChange < 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection_up <- t_DGE_KL %>%
    dplyr::filter(log2FoldChange > 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection <- c(selection_down[["feature"]], selection_up[["feature"]]) %>%
        as.character()

title <- paste0(
    "volcano plot | K. lactis features |\n",
    "size factors estimated with all K. lactis features"
)
subtitle <- paste(
    "points: S. cerevisiae features",
    "| left: up in N",
    "| right: up in SS",
    "| labels: top 5 N and top 5 SS features"
)  # [1] "transcription_SS_vs_N"
p_KL <- plot_volcano(
    table = t_DGE_KL,
    label = all,
    selection = selection,
    label_size = 2.5,
    p_cutoff = 0.05,
    FC_cutoff = 1,
    xlim = c(-14, 14),
    ylim = c(0, 310),
    color = "#448EE2",
    title = title,
    subtitle = subtitle
)
p_KL
# save_volcano(
#     file = "test.pdf",
#     plot = p,
#     width = 2,
#     height = 3
# )

rm(all, selection, selection_up, selection_down, title, subtitle)
```
<br />
<br />

## III. Run analyses of *S.C.* in which all *K.L.* are `controlGenes`
### Perform size-factor estimation
To run analyses using all *K.lactis* features as `controlGenes`, we use a
logical vector (a vector composed of elements with values of either `TRUE` or
`FALSE`) obtained from parsing the `rowRanges` dataframe within the `dds`
object. In essence, we're saying, "Return `TRUE` if the `rowRanges` variable
`genome` has a value of `K_lactis`; otherwise, return `FALSE`." Then,
`BiocGenerics::estimateSizeFactors()` is using the values associated with those
`TRUE`s to isolate the counts for *K. lactis*-specific features, in turn
using those values to calculate size factors.

```{r III. Perform size-factor estimation, results='hide', message=FALSE}
#  Recalculate size factors using K. lactis features as `controlGenes`
dds_SC.ctrl_KL <- BiocGenerics::estimateSizeFactors(
    dds,
    controlGenes = (dds@rowRanges$genome == "K_lactis")
)
dds_SC.ctrl_KL@colData
#  Using all of the K. lactis features as `controlGenes`
# Q N rep1 6.264768
# Q N rep2 7.190513
# Q N rep3 6.525544
# Q SS rep1 0.161401
# Q SS rep2 0.165265
# Q SS rep3 0.140550

#NOTE 1/2 Note the much smaller (SS) and much larger (N) coefficients needed to
#NOTE 2/2 scale the libraries
```
<br />

### Run `DESeq2`
#### Call `DESeq2` using default parameters
Since we've already calculated the size factors, I think we can exclude
*K. lactis* features from our work from here on out.
```{r III. Call DESeq2 using default parameters, results='hide', message=FALSE}
dds_SC.ctrl_KL <- DESeq2::DESeq(
    dds_SC.ctrl_KL[dds_SC.ctrl_KL@rowRanges$genome != "K_lactis", ]
)

#  Check model information
DESeq2::resultsNames(dds_SC.ctrl_KL)[2]
#  [1] "transcription_SS_vs_N"
#+ Thus, the model varies on transcription, SS is the numerator, N is the
#+ denominator
```
<br />

#### Call `DESeq2::results()`
```{r III. Call DESeq2::results(), message=FALSE}
#  Set up necessary parameters for generation of DESeq2 results table
independent_filtering <- TRUE
threshold_p <- 0.05
threshold_lfc <- 0

#  Output a DESeq2 DataFrame object
DGE_unshrunken_DF_SC.ctrl_KL <- DESeq2::results(
    dds_SC.ctrl_KL,
    name = DESeq2::resultsNames(dds_SC.ctrl_KL)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "DataFrame"
)

#  Output a GRanges object, which we can easily add to and convert to other
#+ formats (such as a tibble)
DGE_unshrunken_GR_SC.ctrl_KL <- DESeq2::results(
    dds_SC.ctrl_KL,
    name = DESeq2::resultsNames(dds_SC.ctrl_KL)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "GRanges"
)
DGE_unshrunken_GR_SC.ctrl_KL$feature <-
    MatrixGenerics::rowRanges(dds_SC.ctrl_KL)$feature
DGE_unshrunken_GR_SC.ctrl_KL$type <-
    MatrixGenerics::rowRanges(dds_SC.ctrl_KL)$type
DGE_unshrunken_GR_SC.ctrl_KL$genome <-
    MatrixGenerics::rowRanges(dds_SC.ctrl_KL)$genome

t_DGE_SC.ctrl_KL <- DGE_unshrunken_GR_SC.ctrl_KL %>% dplyr::as_tibble()

rm(independent_filtering, threshold_p, threshold_lfc)
```
<br />

#### Make an MA plot that colors features by independent filtering
```{r III. Make an MA plot: B, message=FALSE}
#  Set up temporary variable 'tbl', which will be passed to ggplot
tbl <- t_DGE_SC.ctrl_KL
tbl <- tbl[with(tbl, order(log2FoldChange)), ]
tbl$threshold <- as.factor(tbl$padj <= 0.05)
tbl$log10baseMean <- ifelse(
    is.infinite(log10(tbl$baseMean)), NA, log10(tbl$baseMean)
)

title <- paste0(
    "MA plot | S. cerevisiae features |\n",
    "size factors estimated with all K. lactis features"
)
subtitle <- paste(
    "points: S. cerevisiae features",
    "| top: up in SS",
    "| bottom: up in N"
)  # [1] "transcription_SS_vs_N"
ggplot(tbl, aes(x = log10baseMean, y = log2FoldChange, colour = threshold)) +
    geom_point(alpha = 0.25, size = 0.5) +
    # geom_hline(aes(yintercept = 0), colour = "#000000", linewidth = 0.25) +
    geom_hline(aes(yintercept = 0), colour = "#000000", size = 0.25) +
    # ylim(c(min(tbl$log2FoldChange), max(tbl$log2FoldChange))) +
    ylim(c(-14, 14)) +
    xlab("log10(mean normalized counts)") +
    ylab("log2(fold change)") +
    scale_colour_discrete(name = "q ≤ 0.05") +
    ggtitle(title, subtitle) +
    theme_slick
#TODO 1/2 Explain and make a decision regarding use of 'size' or 'linewidth'
#TODO 2/2 parameters

#  Create a vector of features that both passed independent filtering (and thus
#+ have an inherently high mean expression) and are not statistically
#+ significant; this vector signifies features that are "stably expressed"
#+ between conditions
tbl$stably_expressed <- ifelse(
    !is.na(tbl$threshold) & tbl$padj > 0.05,
    TRUE,
    FALSE
)
stably_expressed_SC.ctrl_KL <- tbl$stably_expressed[
    !is.na(tbl$stably_expressed)
]

rm(tbl, title, subtitle)
```

#### Make a volcano plot
```{r III. Make a volcano plot, message=FALSE}
all <- t_DGE_SC.ctrl_KL$feature
selection_down <- t_DGE_SC.ctrl_KL %>%
    dplyr::filter(log2FoldChange < 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection_up <- t_DGE_SC.ctrl_KL %>%
    dplyr::filter(log2FoldChange > 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection <- c(selection_down[["feature"]], selection_up[["feature"]]) %>%
        as.character()

title <- paste0(
    "volcano plot | S. cerevisiae features |\n",
    "size factors estimated with all K. lactis features"
)
subtitle <- paste(
    "points: S. cerevisiae features",
    "| left: up in N",
    "| right: up in SS",
    "| labels: top 5 N and top 5 SS features"
)  # [1] "transcription_SS_vs_N"
p_SC.ctrl_KL <- plot_volcano(
    table = t_DGE_SC.ctrl_KL,
    label = all,
    selection = selection,
    label_size = 2.5,
    p_cutoff = 0.05,
    FC_cutoff = 1,
    xlim = c(-14, 14),
    ylim = c(0, 310),
    color = "#A020F0",
    title = title,
    subtitle = subtitle
)
p_SC.ctrl_KL
# save_volcano(
#     file = "test.pdf",
#     plot = p,
#     width = 2,
#     height = 3
# )

rm(all, selection, selection_up, selection_down, title, subtitle)
```
<br />
<br />
