---
title: "work_normalization-etc_rough-draft"
author: "KA"
email: "kalavatt@fredhutch.org"
output: html_notebook
---
<br />

## Prepare data for DGE analyses
### Get situated
#### Set working directory
```{r set working directory, results='hide', message=FALSE}
# p_local <- "/Users/kalavattam/Dropbox/FHCC"  # KrisMac
p_local <- "/Users/kalavatt/projects-etc"  # WorkMac
p_wd <- "2022_transcriptome-construction/results/2023-0215"
setwd(paste(p_local, p_wd, sep = "/"))
getwd()

rm(p_local, p_wd)
```
<br />

#### Load necessary libraries
```{r load libraries, results='hide', message=FALSE, warning=FALSE}
library(DESeq2)
library(EnhancedVolcano)
library(GenomicRanges)
library(IRanges)
library(tidyverse)
```
<br />

#### Initialize necessary functions
```{r load functions, results='hide', message=FALSE}
split_isolate_convert <- function(in_vector, field, column_name) {
    out_df <- in_vector %>%
        stringr::str_split(., c("_")) %>%
        sapply(., "[", field) %>%
        as.data.frame() %>%
        tibble::as_tibble()
    
    colnames(out_df) <- column_name
    
    return(out_df)
}


plot_volcano <- function(
    table, label, selection, label_size, p_cutoff, FC_cutoff,
    xlim, ylim, color, title, subtitle, ...
) {
    # :param table: dataframe of test statistics
    # :param label: character vector of all variable names in param table
    # :param selection: character vector of selected variable names in param
    #                   table
    # :param label_size: size of label font [float]
    # :param p_cutoff: cut-off for statistical significance; a horizontal line
    #                  will be drawn at -log10(pCutoff); p is actually padj
    # :param FC_cutoff: cut-off for absolute log2 fold-change; vertical lines
    #                   will be drawn at the negative and positive values of
    #                   log2FCcutoff
    # :param xlim: limits of the x-axis
    # :param ylim: limits of the y-axis
    # :param color: color of DEGs, e.g., '#52BE9B'
    # :param title: plot title
    # :param subtitle: plot subtitle
    volcano <- EnhancedVolcano::EnhancedVolcano(
        toptable = table,
        lab = label,
        selectLab = selection,
        x = "log2FoldChange",
        y = "padj",
        xlab = "log2(FC)",
        ylab = "-log10(padj)",
        pCutoff = p_cutoff,
        pCutoffCol = "padj",
        FCcutoff = FC_cutoff,
        xlim = xlim,
        ylim = ylim,
        cutoffLineType = "dashed",
        cutoffLineWidth = 0.2,
        pointSize = 1,
        shape = 16,
        colAlpha = 0.25,
        col = c('#D3D3D3', '#D3D3D3', '#D3D3D3', color),
        title = NULL,
        subtitle = NULL,
        caption = NULL,
        borderColour = "#000000",
        borderWidth = 0.2,
        gridlines.major = TRUE,
        gridlines.minor = TRUE,
        axisLabSize = 10,
        labSize = label_size,
        boxedLabels = TRUE,
        parseLabels = TRUE,
        drawConnectors = TRUE,
        widthConnectors = 0.2,
        colConnectors = 'black',
        max.overlaps = Inf
    ) +
        theme_slick_no_legend +
        ggplot2::ggtitle(title, subtitle = subtitle)
    return(volcano)
}


save_volcano <- function(plot, file, width, height) {
    ggplot2::ggsave(
        filename = file,
        plot,
        device = "pdf",
        h = width,
        w = height,
        units = "in"
    )
}


#  Set up custom ggplot2 plot themes ------------------------------------------
theme_slick <- theme_classic() +
    theme(
        panel.grid.major = ggplot2::element_line(linewidth = 0.4),
        panel.grid.minor = ggplot2::element_line(linewidth = 0.2),
        axis.line = ggplot2::element_line(linewidth = 0.2),
        axis.ticks = ggplot2::element_line(linewidth = 0.4),
        axis.text = ggplot2::element_text(color = "black"),
        axis.title.x = ggplot2::element_text(),
        axis.title.y = ggplot2::element_text(),
        plot.title = ggplot2::element_text(),
        text = element_text(family = "")
    )

theme_slick_no_legend <- theme_slick + theme(legend.position = "none")
```
<br />

### Load in and process `featureCounts` table
```{r process featureCounts table, results='hide', message=FALSE}
#  Load in featureCounts table ------------------------------------------------
p_fc <- "outfiles_featureCounts/combined_SC_KL/UT_prim_UMI"
f_fc <- "UT_prim_UMI.featureCounts"
t_fc <- read.table(
    paste(p_fc, f_fc, sep = "/"), header = TRUE, row.names = 1
) %>% 
    tibble::rownames_to_column() %>%
    tibble::as_tibble()

rm(p_fc, f_fc)
#NOTE 1/2 Alison, you can adjust this to point to wherever
#NOTE 2/2 UT_prim_UMI.featureCounts is on your system


#  Clean up tibble column names -----------------------------------------------
colnames(t_fc) <- colnames(t_fc) %>%
    gsub("rowname", "feature_init", .) %>%
    gsub("Chr", "chr", .) %>%
    gsub("Start", "start", .) %>%
    gsub("End", "end", .) %>%
    gsub("Strand", "strand", .) %>%
    gsub("Length", "length", .) %>%
    gsub("bams_renamed\\.UT_prim_UMI\\.", "", .) %>%
    gsub(".UT_prim_UMI.bam", "", .) %>%
    gsub("\\.d", "-d", .) %>%
    gsub("\\.n", "-n", .) %>%
    gsub("aux\\.", "aux-", .) %>%
    gsub("tc\\.", "tc-", .)


#  Order tibble by chromosome names and feature start positions ---------------
chr_SC <- c(
    "I", "II", "III", "IV", "V", "VI", "VII", "VIII", "IX", "X", "XI", "XII",
    "XIII", "XIV", "XV", "XVI", "Mito"
)
chr_KL <- c("A", "B", "C", "D", "E", "F")
chr_order <- c(chr_SC, chr_KL)
t_fc$chr <- t_fc$chr %>% as.factor()
t_fc$chr <- ordered(t_fc$chr, levels = chr_order)

t_fc <- t_fc %>% dplyr::arrange(chr, start)


#  Categorize chromosomes by genome of origin ---------------------------------
t_fc$genome <- ifelse(
    t_fc$chr %in% chr_SC,
    "S_cerevisiae",
    ifelse(
        t_fc$chr %in% chr_KL,
        "K_lactis",
        NA
    )
) %>%
    as.factor()

#  Move the new column 'genome' to a better location in the tibble
t_fc <- t_fc %>% dplyr::relocate("genome", .before = "chr")

#  Check on variable "genome"
levels(t_fc$genome)
t_fc %>%
    dplyr::group_by(genome) %>%
    dplyr::summarize(tally = length(genome))
# K_lactis = 5659, S_cerevisiae = 7507

rm(chr_KL, chr_SC, chr_order)


#  Split and better organize variable 'feature_init' --------------------------
#  Split 'feature_init' into two distinct elements (separated by an underscore)
el_1 <- split_isolate_convert(
    in_vector = t_fc$feature_init,
    field = 1,
    column_name = "feature"
)
el_2 <- split_isolate_convert(
    in_vector = t_fc$feature_init,
    field = 2,
    column_name = "type"
)

#  Append split information to tibble 't_fc'
t_fc <- dplyr::bind_cols(t_fc, el_1, el_2) %>%
    dplyr::relocate(c("feature", "type"), .after = "feature_init")

rm(el_1, el_2)

#  Limit the splitting/reorganization to S. cerevisiae features only (it's
#+ not appropriate for the K. lactis 'feature_init' information given the
#+ organism's naming/classification scheme)
t_fc$feature <- ifelse(
    t_fc$genome == "K_lactis", t_fc$feature_init, t_fc$feature
)
t_fc$type <- ifelse(
    t_fc$genome == "K_lactis", NA, t_fc$type
)

#  Create levels for S. cerevisiae 'type' NAs and K. lactis 'type' NAs, then
#+ factorize variable 'type'
t_fc$type <-  ifelse(
    (t_fc$genome == "S_cerevisiae" & is.na(t_fc$type)),
    "NA_SC",
    ifelse(
        (t_fc$genome == "K_lactis" & is.na(t_fc$type)),
        "NA_KL",
        t_fc$type
    )
) %>%
    as.factor()

# #  Do a quick check of the tibble
# t_fc

#  Check on the split information
levels(t_fc$type)  # 19 levels
t_fc %>%
    dplyr::group_by(type) %>%
    dplyr::summarize(tally = length(type))
#  Quick check:
#+ mRNA-E1 = 6600, mRNA-E2 = 283, NA_KL = 5547, NA_SC = 103, tRNA-E1 = 299,
#+ tRNA-E2 = 60, etc.
```
<br />

### Record tibble's positional information in `GRanges` object
*(`pos_info` object to be used in the DESeq2 processing, post-processing, etc.)*

```{r positional information, results='hide', message=FALSE}
pos_info <- GenomicRanges::GRanges(
    seqnames = t_fc$chr,
    ranges = IRanges::IRanges(t_fc$start, t_fc$end),
    strand = t_fc$strand,
    length = t_fc$length,
    feature = t_fc$feature,
    type = t_fc$type,
    genome = t_fc$genome
)
pos_info
```
<br />
<br />

## Perform normalization and run DGE analyses
### Do prep work with everything
#### Establish table of variables for dds, i.e., a "master" model matrix
- `dds` is `DESeqDataSet` object
- variables for `dds` are
    + `strain`
    + `state`
    + `time`
    + `kit` *(`tcn` for Tecan, `ovn` for Ovation)*
    + `transcription` *(`N` for nascent, `SS` for steady state)*
    + `auxin`
    + `timecourse`
    + `replicate`

```{r make master model matrix, results='hide', message=FALSE}
#  Columns 7 through to the last column are samples; get them into a vector
samples <- colnames(t_fc)[10:length(colnames(t_fc))]

#  Convert vector to list by splitting each element at underscores
samples <- stringr::str_split(samples, "_")

#  Convert the list to a df, transpose it, then convert it to a tibble
samples <- samples %>%
    as.data.frame(
        .,
        col.names = c(seq(1, 55)),
        row.names = c(
            "strain", "state", "time", "kit", "transcription", "auxin",
            "timecourse", "replicate"
        )
    ) %>%
    t() %>%
    tibble::as_tibble()

#  Add a "keys" variable for quickly accessing combinations of variable values
keys <- vector(mode = "character")
for(i in seq(1, nrow(samples))) {
    # i <- 1
    keys[i] <- paste(
        samples[i, 1], samples[i, 2], samples[i, 3], samples[i, 4],
        samples[i, 5], samples[i, 6], samples[i, 7], samples[i, 8],
        sep = "_"
    )
}
keys <- keys %>% as.data.frame()
colnames(keys) <- "keys"

samples <- dplyr::bind_cols(samples, keys) %>%
    dplyr::relocate("keys", .before = "strain")

rm(i)

#  Convert all columns to data type 'factor' (helps with running
#+ DESeq2::DESeqDataSetFromMatrix())
samples[sapply(samples, is.character)] <- lapply(
    samples[sapply(samples, is.character)], as.factor
)

#  How does it look?
samples
```
<br />

### Do prep work with `WT_{G1,Q}_day7_ovn_N_aux-F_tc-F_rep{1,2}`
*(i.e., the wild-type G1 and Q Ovation 4tU-seq datasets)*
<br />

#### Make the counts matrix
```{r make counts matrix, results='hide', message=FALSE}
datasets <- c(
    "WT_G1_day1_ovn_N_aux-F_tc-F_rep1",
    "WT_G1_day1_ovn_N_aux-F_tc-F_rep2",
    "WT_Q_day7_ovn_N_aux-F_tc-F_rep1",
    "WT_Q_day7_ovn_N_aux-F_tc-F_rep2"
)
counts_data <- t_fc[, colnames(t_fc) %in% datasets] %>%
    as.data.frame()

#  How do things look?
counts_data
```
<br />

#### Make the model matrix
```{r make model matrix, results='hide', message=FALSE}
#  Use the "keys" column to isolate WT G1 and Q Ovation 4tU-seq datasets
#REMEMBER variable `datasets` is initialized in the preceding chunk 
col_data <- samples[samples$keys %in% datasets, ] %>%
    as.data.frame() %>%
    tibble::column_to_rownames(., var = "keys")

#  How do things look?
col_data
```
<br />

#### Make the `DESeqDataSet`, `dds`
- `counts_data` for the `featureCount` tallies
- `col_data` for modeling
- `pos_info` for subsetting, etc.

```{r make dds, results='hide', message=FALSE}
dds <- DESeq2::DESeqDataSetFromMatrix(
    countData = counts_data,
    colData = col_data,
    design = ~ state,
    rowRanges = pos_info
)

#  Make a back-up of the DESeqDataSet object
bak.dds <- dds

#  How do things look?
# dds %>% BiocGenerics::counts() %>% head()
# dds@rowRanges
# dds@design
# dds@assays

#TODO 1/2 Mess around with complicated models for Q vs G1 w/r/t N vs SS later
#TODO 2/2 e.g., `design = ~ state + transcription + state:transcription`
```
<br />

#### Prefilter `dds`
*(Probably don't need to do this; we can consider it for later)*
```{r prefilter dds, results='hide', message=FALSE}
# threshold <- 0
# dds_filt <- dds[rowSums(BiocGenerics::counts(dds)) >= threshold, ]
#
#  Breakdown
#    0 13166
#   10 11893
#   50 11049
#  100 10645
#  500 8601
# 1000 6738
#
# rm(threshold, dds_filt)
```
<br />
<br />

## Run analyses with *S. cerevisae* features only
### Perform size-factor estimation
Here, we subset out the *K. lactis* features and are thus using only
*S. cerevisiae* features in the size-factor estimation. No control genes are
used.

```{r do SC size-factor estimation, results='hide', message=FALSE}
dds_SC <- BiocGenerics::estimateSizeFactors(
    dds[dds@rowRanges$genome != "K_lactis", ]
)
dds_SC@colData
#  Not using controlGenes; calculate size factors with S. cerevisiae features
# G1 rep1 1.144887
# G1 rep2 1.166906
#  Q rep1 0.784023
#  Q rep2 0.970161
```
<br />

#### Do rlog normalization
On the `blind` option: "`logical`, whether to blind the transformation to the
experimental design. `blind = TRUE` should be used for comparing samples in a
manner unbiased by prior information on samples, for example to perform sample
QA (quality assurance). `blind = FALSE` should be used for transforming data
for downstream analysis, where the full use of the design information should be
made. `blind = FALSE` will skip re-estimation of the dispersion trend if this
has already been calculated. If many of genes have large differences in counts
due to the experimental design, it is important to set `blind = FALSE` for
downstream analysis."

```{r rlog normalization, results='hide', message=FALSE}
# rlog_F_SC <- DESeq2::rlog(
#     dds_SC,
#     blind = FALSE
# ) %>%
#     SummarizedExperiment::assay() %>%
#     tibble::as_tibble() %>%
#     dplyr::bind_cols(
#         .,
#         tibble::as_tibble(dds_SC@rowRanges)
#     )
# 
# rlog_T_SC <- DESeq2::rlog(
#     dds_SC,
#     blind = TRUE
# ) %>%
#     SummarizedExperiment::assay() %>%
#     tibble::as_tibble() %>%
#     dplyr::bind_cols(
#         .,
#         tibble::as_tibble(dds_SC@rowRanges)
#     )
# 
# #  How does rlog_T look?
# rlog_T_SC
```
<br />

### Run `DESeq2`
#### Call `DESeq2` using default parameters
```{r call deseq2, results='hide', message=FALSE}
dds_SC <- DESeq2::DESeq(dds_SC)
```
<br />

#### Call `DESeq2::results()`
We need two kinds of output from `DESeq2::results()`: A `DataFrame` object and
a `GRanges` object; we need our data in two different data formats for use with
different functions in downstream analyses.
```{r call results, results='hide', message=FALSE}
independent_filtering <- TRUE
threshold_p <- 0.05
threshold_lfc <- 0

#  Output a DESeq2 DataFrame object
DGE_unshrunken_DF_SC <- DESeq2::results(
    dds_SC,
    name = DESeq2::resultsNames(dds_SC)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "DataFrame"
)

#  Output a GRanges object, which we can easily add to and convert to other
#+ formats (such as a tibble)
DGE_unshrunken_GR_SC <- DESeq2::results(
    dds_SC,
    name = DESeq2::resultsNames(dds_SC)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "GRanges"
)
DGE_unshrunken_GR_SC$feature <- MatrixGenerics::rowRanges(dds_SC)$feature
DGE_unshrunken_GR_SC$type <- MatrixGenerics::rowRanges(dds_SC)$type
DGE_unshrunken_GR_SC$genome <- MatrixGenerics::rowRanges(dds_SC)$genome

t_DGE_SC <- DGE_unshrunken_GR_SC %>%
    dplyr::as_tibble() %>%
    dplyr::rename(chr = seqnames)

rm(independent_filtering, threshold_p, threshold_lfc)
```
<br />

#### Make an MA plot
```{r}
#REMEMBER 1/2 No `controlGenes` used; calculation of size factors with only S.
#REMEMBER 2/2 cerevisiae features
DESeq2::plotMA(
    DGE_unshrunken_DF_SC,
    colNonSig = "#D3D3D390",
    colSig = "#52BE9B90",
    colLine = "#82858790",
    ylim = c(-14, 14),
    frame = FALSE,
)
#  Plot of only S. cerevisiae points (K. lactis points already gone)
#+ - top: up in Q
#+ - bottom: up in G1

#  The data are centered nicely on 0, suggesting the normalization we used is
#+ appropriate
```
<br />

#### Make an MA plot that colors features that passed independent filtering
```{r}
tbl <- t_DGE_KL
tbl <- tbl[with(tbl, order(log2FoldChange)), ]
tbl$threshold <- as.factor(tbl$padj <= 0.05)
tbl$log10baseMean <- ifelse(
    is.infinite(log10(tbl$baseMean)), NA, log10(tbl$baseMean)
)

ggplot(tbl, aes(x = log10baseMean, y = log2FoldChange, colour = threshold)) +
    geom_point(alpha = 0.25, size = 0.5) +
    geom_hline(aes(yintercept = 0), colour = "#000000", linewidth = 0.25) +
    # ylim(c(min(tbl$log2FoldChange), max(tbl$log2FoldChange))) +
    ylim(c(-6, 6)) +
    xlab("log10(mean normalized counts)") +
    ylab("log2(FC)") +
    theme(
        axis.title.x = element_text(face = "bold", size = 15),
        axis.text.x = element_text(face = "bold", size = 12)
    ) +
    theme(
        axis.title.y = element_text(face = "bold", size = 15),
        axis.text.y = element_text(face = "bold", size = 12)
    ) +
    scale_colour_discrete(name = "q ≤ 0.05") +
    theme(legend.title = element_text(face = "bold", size = 15)) +
    theme(legend.text = element_text(size = 14)) +
    theme_slick
```
<br />

#### Draw an MA plot that colors features that passed independent filtering
```{r}
tbl <- t_DGE_KL
tbl <- tbl[with(tbl, order(log2FoldChange)), ]
tbl$threshold <- as.factor(tbl$padj <= 0.05)
tbl$log10baseMean <- ifelse(
    is.infinite(log10(tbl$baseMean)), NA, log10(tbl$baseMean)
)

ggplot(tbl, aes(x = log10baseMean, y = log2FoldChange, colour = threshold)) +
    geom_point(alpha = 0.25, size = 0.5) +
    geom_hline(aes(yintercept = 0), colour = "#000000", linewidth = 0.25) +
    # ylim(c(min(tbl$log2FoldChange), max(tbl$log2FoldChange))) +
    ylim(c(-6, 6)) +
    xlab("log10(mean normalized counts)") +
    ylab("log2(FC)") +
    theme(
        axis.title.x = element_text(face = "bold", size = 15),
        axis.text.x = element_text(face = "bold", size = 12)
    ) +
    theme(
        axis.title.y = element_text(face = "bold", size = 15),
        axis.text.y = element_text(face = "bold", size = 12)
    ) +
    scale_colour_discrete(name = "q ≤ 0.05") +
    theme(legend.title = element_text(face = "bold", size = 15)) +
    theme(legend.text = element_text(size = 14)) +
    theme_slick

#  Create a list of features that passed independent filtering (and thus have
#+ an inherently high mean expression) and are not statistically significant
#+ (when DESeq2 is called with default settings)
tbl$stably_expressed <- ifelse(
    !is.na(tbl$threshold), tbl$feature, NA_character_
)
stably_expressed_KL <- tbl$stably_expressed[!is.na(tbl$stably_expressed)]

rm(tbl)
```
<br />

#### Make a volcano plot
```{r}
#  Identify and isolate the top 5 upregulated features and the top 5 down-
#+ regulated features
all <- t_DGE_SC$feature
selection_down <- t_DGE_SC %>%
    dplyr::filter(log2FoldChange < 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection_up <- t_DGE_SC %>%
    dplyr::filter(log2FoldChange > 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection <- c(selection_down[["feature"]], selection_up[["feature"]]) %>%
        as.character()

title <- "Size factors estimated w/all SC"
subtitle <- "points: SC; left: up G1; right: up Q"
p_SC <- plot_volcano(
    table = t_DGE_SC,
    label = all,
    selection = selection,
    label_size = 2.5,
    p_cutoff = 0.05,
    FC_cutoff = 1,
    xlim = c(-14, 14),
    ylim = c(0, 310),
    color = "#52BE9B",
    title = title,
    subtitle = subtitle
)
p_SC
# save_volcano(
#     file = "test.pdf",
#     plot = p,
#     width = 2,
#     height = 3
# )

rm(all, selection, selection_up, selection_down)
```
<br />
<br />

## Run analyses with *K. lactis* as `controlGenes`
### Perform size-factor estimation
Here, we subset out the *K. lactis* features and are thus using only
*S. cerevisiae* features in the size-factor estimation. No control genes are
used.

```{r size-factor estimation, results='hide', message=FALSE}
dds_KL <- BiocGenerics::estimateSizeFactors(
    dds[dds@rowRanges$genome != "S_cerevisiae", ]
)
dds_KL@colData
#  Not using controlGenes; calculate size factors with S. cerevisiae features
# G1 rep1 0.174943
# G1 rep2 0.339968
#  Q rep1 4.922491
#  Q rep2 3.450797
```
<br />

#### Do rlog normalization
```{r rlog normalization, results='hide', message=FALSE}
# rlog_F_KL <- DESeq2::rlog(
#     dds_KL,
#     blind = FALSE
# ) %>%
#     SummarizedExperiment::assay() %>%
#     tibble::as_tibble() %>%
#     dplyr::bind_cols(
#         .,
#         tibble::as_tibble(dds_KL@rowRanges)
#     )
# 
# rlog_T_KL <- DESeq2::rlog(
#     dds_KL,
#     blind = TRUE
# ) %>%
#     SummarizedExperiment::assay() %>%
#     tibble::as_tibble() %>%
#     dplyr::bind_cols(
#         .,
#         tibble::as_tibble(dds@rowRanges)
#     )
# 
# #  How does rlog_T look?
# rlog_T_KL
```
<br />

### Run `DESeq2`
#### Call `DESeq2` using default parameters
```{r call deseq2, results='hide', message=FALSE}
dds_KL <- DESeq2::DESeq(dds_KL)
```
<br />

#### Call `DESeq2::results()`
```{r call results, results='hide', message=FALSE}
independent_filtering <- TRUE
threshold_p <- 0.05
threshold_lfc <- 0

#  Output a DESeq2 DataFrame object
DGE_unshrunken_DF_KL <- DESeq2::results(
    dds_KL,
    name = DESeq2::resultsNames(dds_KL)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "DataFrame"
)

#  Output a GRanges object, which we can easily add to and convert to other
#+ formats (such as a tibble)
DGE_unshrunken_GR_KL <- DESeq2::results(
    dds_KL,
    name = DESeq2::resultsNames(dds_KL)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "GRanges"
)
DGE_unshrunken_GR_KL$feature <- MatrixGenerics::rowRanges(dds_KL)$feature
DGE_unshrunken_GR_KL$type <- MatrixGenerics::rowRanges(dds_KL)$type
DGE_unshrunken_GR_KL$genome <- MatrixGenerics::rowRanges(dds_KL)$genome

t_DGE_KL <- DGE_unshrunken_GR_KL %>%
    dplyr::as_tibble() %>%
    dplyr::rename(chr = seqnames)

rm(independent_filtering, threshold_p, threshold_lfc)
```
<br />

#### Make an MA plot
```{r}
#REMEMBER 1/2 No `controlGenes` used; calculation of size factors with only S.
#REMEMBER 2/2 cerevisiae features
DESeq2::plotMA(
    DGE_unshrunken_DF_KL,
    colNonSig = "#D3D3D390",
    colSig = "#448EE290",
    colLine = "#82858790",
    ylim = c(-14, 14),
    frame = FALSE,
)
#  Plot of only K. lactis points (S. cerevisiae points already gone)
#+ - top: up in Q
#+ - bottom: up in G1

#  The data seem to be centered on 0 as expression levels increase; lowly
#+ expressed features skew to negative integers
```
<br />

#### Draw an MA plot that colors features that passed independent filtering
```{r}
tbl <- t_DGE_KL
tbl <- tbl[with(tbl, order(log2FoldChange)), ]
tbl$threshold <- as.factor(tbl$padj <= 0.05)
tbl$log10baseMean <- ifelse(
    is.infinite(log10(tbl$baseMean)), NA, log10(tbl$baseMean)
)

ggplot(tbl, aes(x = log10baseMean, y = log2FoldChange, colour = threshold)) +
    geom_point(alpha = 0.25, size = 0.5) +
    geom_hline(aes(yintercept = 0), colour = "#000000", linewidth = 0.25) +
    # ylim(c(min(tbl$log2FoldChange), max(tbl$log2FoldChange))) +
    ylim(c(-6, 6)) +
    xlab("log10(mean normalized counts)") +
    ylab("log2(FC)") +
    theme(
        axis.title.x = element_text(face = "bold", size = 15),
        axis.text.x = element_text(face = "bold", size = 12)
    ) +
    theme(
        axis.title.y = element_text(face = "bold", size = 15),
        axis.text.y = element_text(face = "bold", size = 12)
    ) +
    scale_colour_discrete(name = "q ≤ 0.05") +
    theme(legend.title = element_text(face = "bold", size = 15)) +
    theme(legend.text = element_text(size = 14)) +
    theme_slick

#  Create a list of features that passed independent filtering (and thus have
#+ an inherently high mean expression) and are not statistically significant
#+ (when DESeq2 is called with default settings)
tbl$stably_expressed <- ifelse(
    !is.na(tbl$threshold), tbl$feature, NA_character_
)
stably_expressed_KL <- tbl$stably_expressed[!is.na(tbl$stably_expressed)]

rm(tbl)
```
<br />

#### Make a volcano plot
```{r}
all <- t_DGE_KL$feature
selection_down <- t_DGE_KL %>%
    dplyr::filter(log2FoldChange < 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection_up <- t_DGE_KL %>%
    dplyr::filter(log2FoldChange > 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection <- c(selection_down[["feature"]], selection_up[["feature"]]) %>%
        as.character()

title <- "Size factors estimated w/all KL"
subtitle <- "points: KL; left: up G1; right: up Q"
p_KL <- plot_volcano(
    table = t_DGE_KL,
    label = all,
    selection = NULL,
    label_size = 2.5,
    p_cutoff = 0.05,
    FC_cutoff = 1,
    xlim = c(-3, 3),
    ylim = c(0, 20),
    color = "#52BE9B",
    title = title,
    subtitle = subtitle
)
p_KL
# save_volcano(
#     file = "test.pdf",
#     plot = p,
#     width = 2,
#     height = 3
# )
```
<br />
<br />



## Run analyses with *K. lactis* as `controlGenes`
### Perform normalization: `WT_{G1,Q}_day7_ovn_N_aux-F_tc-F_rep{1,2}`
Here, we're using a logical vector (vector composed of elements with values of
either `TRUE` or `FALSE`) obtained from parsing the `rowRanges` data frame
within the `dds` object. Specifically, we're saying, "Return `TRUE` if the
`rowRanges` variable `genome` has a value of `K_lactis`; otherwise, return
`FALSE`." `BiocGenerics::estimateSizeFactors()` is using the values associated
with those `TRUE`s to isolate the counts for *K. lactis*-specific features  
and then use those values to calculate the size factors.

```{r}
#  Load the backup DESeqDataSet
dds <- bak.dds

#  Recalculate size factors using K. lactis features as `controlGenes`
dds <- BiocGenerics::estimateSizeFactors(
    dds,
    controlGenes = (dds@rowRanges$genome == "K_lactis")
)
dds@colData
#  Using all of the K. lactis features as controlGenes
# G1 rep1 0.174943
# G1 rep2 0.339968
#  Q rep1 4.922491
#  Q rep2 3.450797

#NOTE Note the much smaller (G1) and much larger (Q) coefficients
```
<br />

#### Do rlog normalization: `WT_{G1,Q}_day7_ovn_N_aux-F_tc-F_rep{1,2}`
Here, we exclude the *K. lactis* data prior to doing the `rlog` normalizations.

```{r}
#  Perform rlog normalization after subsetting out the K. lactis genes
rlog_F <- DESeq2::rlog(
    dds_KL[dds_KL@rowRanges$genome != "K_lactis", ],
    blind = FALSE
) %>%
    SummarizedExperiment::assay() %>%
    tibble::as_tibble() %>%
    dplyr::bind_cols(
        .,
        tibble::as_tibble(dds_KL@rowRanges)[dds_KL@rowRanges$genome != "K_lactis", ]
    )

rlog_T <- DESeq2::rlog(
    dds_KL[dds_KL@rowRanges$genome != "K_lactis", ],
    blind = TRUE
) %>%
    SummarizedExperiment::assay() %>%
    tibble::as_tibble() %>%
    dplyr::bind_cols(
        .,
        tibble::as_tibble(dds_KL@rowRanges)[dds_KL@rowRanges$genome != "K_lactis", ]
    )

#  How does rlog_T look?
rlog_T
#  (Results are the same as above)
```
<br />

### Run `DESeq2`: `WT_{G1,Q}_day7_ovn_N_aux-F_tc-F_rep{1,2}`
#### Call `DESeq2` using default parameters
Since we've already calculated the size factors, I think we can exclude
*K. lactis* features from our work from here on out.
```{r}
dds_KL <- DESeq2::DESeq(dds_KL[dds_KL@rowRanges$genome != "K_lactis", ])
# using pre-existing size factors
# estimating dispersions
# gene-wise dispersion estimates
# mean-dispersion relationship
# final dispersion estimates
# fitting model and testing
```
<br />

#### Call `DESeq2::results()`
```{r}
independent_filtering <- TRUE
threshold_p <- 0.05
threshold_lfc <- 0

#  Output a DESeq2 DataFrame object
DGE_unshrunken_DF <- DESeq2::results(
    dds,
    name = DESeq2::resultsNames(dds)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "DataFrame"
)

#  Output a GRanges object, which we can easily add to and convert to other
#+ formats (such as a tibble)
DGE_unshrunken_GR <- DESeq2::results(
    dds,
    name = DESeq2::resultsNames(dds)[2],
    independentFiltering = independent_filtering,
    alpha = threshold_p,
    lfcThreshold = threshold_lfc,
    format = "GRanges"
)
DGE_unshrunken_GR$feature <- MatrixGenerics::rowRanges(dds)$feature
DGE_unshrunken_GR$type <- MatrixGenerics::rowRanges(dds)$type
DGE_unshrunken_GR$genome <- MatrixGenerics::rowRanges(dds)$genome

t_DGE <- DGE_unshrunken_GR %>% dplyr::as_tibble()

rm(independent_filtering, threshold_p, threshold_lfc)
```
<br />

#### Make an MA plot
Notice that the data are no longer centered on `log fold change = 0` when we
use all *K. lactis* features as `controlGenes`. The data seem to be centered on
approximately `log fold change = 5`.
```{r}
BiocGenerics::plotMA(DGE_unshrunken_DF)
#  Plot of only S. cerevisiae points
#+ - top: up in Q
#+ - bottom: up in G1
```


#### Make a volcano plot
```{r}
#  Make the plot
all <- t_DGE$feature
selection_down <- t_DGE %>%
    dplyr::filter(log2FoldChange < 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection_up <- t_DGE %>%
    dplyr::filter(log2FoldChange > 0) %>%
    dplyr::arrange(padj) %>%
    dplyr::slice(1:5)
selection <- c(selection_down[["feature"]], selection_up[["feature"]]) %>%
        as.character()

title <- "Size factors est. w/KL. Points: SC; left: up G1; right: up Q."
p <- plot_volcano(
    table = t_DGE,
    label = all,
    selection = selection,
    label_size = 2.5,
    p_cutoff = 0.05,
    FC_cutoff = 1,
    xlim = c(-14, 14),
    ylim = c(0, 310),
    color = "#DC267F",
    title = title
)
p
# save_volcano(
#     file = "test.pdf",
#     plot = p,
#     width = 2,
#     height = 3
# )
```
